Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-05-28 extensions.tex
Filled out extensions section.

This section describes how three example monitoring schemes map to the proposed
architecture.

@@

@@

@@

@@

>>>>

% Monitoring operations
\begin{table*}[tb]
  \begin{center}
    \begin{small}
    \input{tabs/monitor}
    \end{small}
    \caption{Monitoring operations for BC, UMC, and DIFT.}
    \label{tab:extensions.monitor}
    \vspace{-0.2in}
  \end{center}
\end{table*}

% Filtering operations for clean flags
\begin{table*}[tb]
  \begin{center}
    \begin{small}
    \input{tabs/filter}
    \end{small}
    \caption{Filtering conditions and operations for clean flags.}
    \label{tab:extensions.filter}
    \vspace{-0.2in}
  \end{center}
\end{table*}

% Filtering operations for invalid flags
\begin{table*}[tb]
  \begin{center}
    \begin{small}
    \input{tabs/drop}
    \end{small}
    \caption{Filtering conditions and operations performed for a dropped or filtered event for invalid flags.}
    \label{tab:extensions.drop}
    \vspace{-0.2in}
  \end{center}
\end{table*}

In this section, we show examples of how our architecture can be applied to
three specific monitoring schemes: array bounds check (BC), uninitialized
memory check (UMC), and dynamic information flow tracking (DIFT).
Table~\ref{tab:extensions.monitor} shows a summary of the monitoring operations for these
monitoring schemes. Table~\ref{tab:extensions.filter} summarizes when when
monitoring events can be filtered out due to clean metadata and what metadata
needs to be marked as clean on a filtered event.
Table~\ref{tab:extensions.drop} shows the analogous information for the valid
flags. In this case, the operation listed is also the one performed when a
monitoring event is dropped due to insufficient slack. Note that for simplicity
the policies for clean and invalid flags are shown and discussed separately
here. However, the dataflow engine actually operates on pairs of these flags
using a combined policy.

@@

As described in Section~\ref{sec:monitoring}, array bounds check is a
monitoring scheme that attempts to detect buffer overflows where memory
accesses go beyond the boundaries of an array.
Buffer overflows can cause unreliable program execution or create openings for
malicious attackers. One way to detect buffer overflows is a base and bounds
approach \cite{hardbound-asplos08}. In this approach, metadata is
associated with an array pointer when it is allocated that includes the base
(start) address and bound (end) address of the corresponding array. On ALU
instructions, this base and bound metadata is copied to the destination
registers.  On store instructions, the metadata for the source register is
copied to the metadata for the store address. Similarly, on load instructions,
the metadata value for the load address is copied to the destination register's
metadata. In addition, on load and store instructions, the memory address is
checked to be within the base and bounds metadata. If it is not, then an
out-of-bounds error is detected. These operations are summarized in
Table~\ref{tab:extensions.monitor}.

In terms of filtering using clean flags (Table~\ref{tab:extensions.filter}),
array allocations are never filtered since these will always involve non-clean
metadata (i.e., setting new base and bounds). For ALU operations, if both
source operands are clean, then the event can be filtered by setting the
destination register as clean. For load and store, only a single source needs
to be checked to be clean in order for the event to be filtered. Similarly, the
destination register or memory location needs to be marked as clean if the
event is filtered.

The invalid flags for BC operate similarly to the clean flags
(Table~\ref{tab:extensions.drop}). When ALU, load, or store events are dropped or filtered,
the destination register or memory location is marked as invalid. In addition,
we need to handle the possibility of dropping an array allocation. In this
case, the array pointer is marked as invalid. In terms of filtering out invalid
events, for an ALU event, if either of the source registers is invalid then the
event can be filtered. For load and store instructions, if the single source
operand is invalid, then the event can be filtered.

@@

Uninitialized memory check is a monitoring scheme that checks that memory is
written to before it is read from. This is done by forwarding all load and
store instructions to the monitoring core. For every byte of memory, the
monitor keeps one bit of metadata. On a store to a memory location, the monitor
marks the corresponding metadata bit to indicate that the memory location has
been initialized. On a load, the monitor checks that the corresponding metadata
bit has been previously marked as initialized.

Rather than using the clean flags to mirror the initialization metadata that
UMC stores, we use the clean flags as a coarse-grained version of the UMC
metadata in order to more efficiently filter out monitoring operations. We use
a clean flag per word (four bytes) of memory. A clean flag is set to dirty only
if all four of the corresponding memory bytes have been marked as initialized.
On a load, if the clean flag is marked as dirty, then we can filter the event
since we know that all possible corresponding location memory locations have
been marked as initialized. If the flag is marked as clean, then that indicates
that at least one of the corresponding memory bytes has not been initialized
and the monitoring core must perform a further check. Store instructions are
never filtered out.

On a dropped store instruction, the destination memory location is marked as
invalid. On load instructions, we can filter them out if the source memory
location has been invalidated. In this case, we do not have to perform any
further operation since on load instructions, UMC only performs a check
operation.

@@

Dynamic information flow tracking (DIFT) \cite{dift-asplos04} is a security
monitoring scheme that seeks to detect when information from untrusted sources
is used to affect the control flow.  In its simplest form, DIFT keeps a 1-bit
metadata tag for each memory location and each register, indicating tainted or
untainted. When data is read from an untrusted source, it is marked as tainted.
As this data is used for other operations, the results of these operations are
also marked as tainted. On an indirect jump, the taint bit of the register used
is checked. If the register is found to be tainted, then an error is detected.
We focus on a more heavy-weight approach that also allows some information of
where flows originate from by using a multi-bit (32-bits in our implementation)
metadata tag. This can be used to assist debugging in the case of an error. In
this case, when data is marked from an untrusted source, the software specifies
an identifier to set the metadata tag. This multi-bit tag is propagated on
load, store, and ALU operations. If two different tags are used as sources to
an ALU operation, then a new identifier is calculated based on the sources.
These monitoring operations are summarized in
Table~\ref{tab:extensions.monitor}.

Table~\ref{tab:extensions.filter} shows how the clean flags are used for DIFT.
The rules for ALU, load, and store instructions are same as for array bounds
check. Similar to array allocations, for DIFT, software setting of metadata
tags is never filtered out due to clean metadata. On indirect jump
instructions, if the source register is clean, then the event can be filtered
out as no error has occurred.

Again, the invalid flags for DIFT operate similar to their operation for BC
(Table~\ref{tab:extensions.drop}. In addition to operations for invalidating
destination operands for ALU, load, and store instructions, the destination
memory location is also invalidated when set tag operations dropped. Indirect
jump events are filtered out if the source register is marked as invalid.
However, no further metadata needs to be marked as invalid in this case.

@@

In this section, we have discussed how our dataflow engine can be applied to
three example monitoring schemes to enable reduced and adjustable overheads.
However, our approach can be generally applied to many monitoring schemes. This
is possible because the use of clean and invalidation flags is largely agnostic
to the semantics of the specific monitoring scheme. That is, the dataflow
engine must be configured based on the dataflow of the monitoring scheme, but
it does not need detailed information of what the monitoring scheme does.

We do note some limitations to the proposed approach. Previous work
\cite{fade-hpca14} that has looked at filtering in order to reduce monitoring
overheads has also filtered out redundant updates where the source metadata
matches the destination metadata. Since we do not read the actual metadata, our
architecture is not able to identify these redundant updates. However, by using
only 2-bit flags, the bandwidth need of our dataflow engine is reduced.  
In addition, we note that filtering use clean flags works well when a large
portion of instructions are not relevant to the monitoring scheme (e.g., array
bounds check only cares about array pointers). However, if most instruction
have dirty metadata, then very little filtering will be possible.

In terms of using invalidation to perform partial monitoring, this 
approach works best for monitoring schemes where there are many pieces of
independent metadata. Since marking a piece of metadata as invalid will cause
future metadata with dependences to also be marked as invalid, schemes with
large metadata dependency chains can quickly see large amounts of metadata
become invalid.

Another limitation is that the dataflow engine, as described, only marks one
flag location on each monitoring event. Similarly, the dataflow engine can only
read in two sets of flags in order to determine whether an event can be
filtered or not. This was done because we found that for many monitoring
schemes, the monitoring operations corresponded to program instructions, which
read up to two operands and write to one destination.  For monitoring schemes
which update multiple pieces of metadata on a monitoring event, it may be
adequate to associate a clean and invalid flag with a single piece.  For
example, a scheme may use a data structure that includes several pieces of
related metadata. On a drop, one of these pieces could be marked as invalid in
order to indicate that the entire structure is invalid.  Similarly, marking one
specified piece as clean may indicate that the entire data structure's metadata
is clean.  Alternatively, the read and write capabilities of the dataflow
engine could be expanded to support these more complex monitoring schemes.

### 2014-05-28 related.tex
Filled out extensions section.

% Filtering
Hardbound filters out non-pointer data \cite{hardbound-asplos08}. FlexiTaint filters out some monitoring
and performs it in hardware. FADE \cite{fade-hpca14} provides a more general filtering engine. It
filters out redundant updates and clean checks by reading metadata values. Our
dataflow engine does not require reading the actual metadata values which may
be large. We do not handle filtering redundant updates, but we do filter out
clean checks as well as clean (i.e., uninitialized) propagations. We are able
to do this because the dataflow engine can also propagate the uninitialized
information. In addition, our dataflow engine can be used to provide adjustable
overheads without false positives which is not a capability described by FADE.

@@

% Limited Monitoring
In terms of controlling overheads, the Quality Virtual Machine (QVM) is a
modification of the Java Virtual Machine (JVM) that supports run-time
monitoring with controllable overheads \cite{qvm-oopsla08}. Similarly, Huang et
al. created a framework for controlling the overheads of software-based
monitoring \cite{huang-sttt12}. These projects are targeting a similar problem
to the one we address in this paper. However, both projects are focused on
software-based monitoring and thus their mechanisms modify the software
monitoring framework in order to enable or disable monitoring. Instead, our
architecture provides a general hardware mechanism to control overheads of
parallel monitoring architectures. In addition, these works prevented
false positives by enabling and disabling monitoring using a source-dropping
policy. Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
monitoring in hard real-time systems. Their architecture performs last-minute
dropping in order to guarantee strict deadlines. Our system allows for a
trade-off between choosing source-dropping, last-minute dropping, or min-point
dropping depending on the needs of the designer or user which has not been
explored by these previous works.

>>>>

% Filtering
FADE \cite{fade-hpca14} provides a more general filtering engine. It
filters out redundant updates and clean checks by reading metadata values. Our
dataflow engine does not require reading the actual metadata values which may
be large. We do not handle filtering redundant updates, but we do filter out
clean checks as well as clean (i.e., uninitialized) propagations. We are able
to do this because the dataflow engine can also propagate the uninitialized
information. In addition, our dataflow engine can be used to provide adjustable
overheads without false positives which is not a capability described by FADE.

@@

% Limited Monitoring
In terms of controlling overheads, the Quality Virtual Machine (QVM) is a
modification of the Java Virtual Machine (JVM) that supports run-time
monitoring with controllable overheads \cite{qvm-oopsla08}. Similarly, Huang et
al. created a framework for controlling the overheads of software-based
monitoring \cite{huang-sttt12}. These projects are targeting a similar problem
to the one we address in this paper. However, both projects are focused on
software-based monitoring and thus their mechanisms modify the software
monitoring framework in order to enable or disable monitoring. Instead, our
architecture provides a general hardware mechanism to control overheads of
parallel monitoring architectures. In addition, these works prevented
false positives by enabling and disabling monitoring using a source-dropping
policy. Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
monitoring in hard real-time systems. Their architecture performs all-event 
dropping in order to guarantee strict deadlines. Our system allows for a
trade-off between choosing source-event dropping, all-event dropping, or min-event
dropping depending on the needs of the designer or user which has not been
explored by these previous works.
